Finite-size scaling of free energy in the dimer model on a hexagonal domain

### 2019-03-31 AffineLieAlgebras/MQFT2018/paper.tex
First attempt to fix the result

\section{The computation of the free energy asymptotic expansion}
\label{sec:free-energy-scaling}
First we substitute MacMahon formula \eqref{eq:12} into the free energy definition \eqref{eq:17} and
obtain
\begin{equation}
  \label{eq:20}
    %-\frac{f}{T}
    f=-\frac{1}{V}\ln Z =- \sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \frac{1}{V}
  \ln\left(\frac{1-e^{-\varepsilon (i+j+k)} e^{\varepsilon}}{1-e^{-\varepsilon (i+j+k)} e^{2\varepsilon}}\right)
\end{equation}
Expanding the exponents $e^{\varepsilon}, e^{2\varepsilon}$ and logarithms in powers of $\varepsilon$ we obtain

\begin{multline}
  \label{eq:4}
  %\frac{f}{T}
  f=- \sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \frac{1}{V}\left\{
    \varepsilon g(i\varepsilon,j\varepsilon,k\varepsilon)+
    \frac{3}{2}\varepsilon^{2}\left[g(i\varepsilon,j\varepsilon,k\varepsilon)+g(i\varepsilon,j\varepsilon,k\varepsilon)^{2}\right]\right.\\
  \left.+\frac{7}{6}\varepsilon^{3}\left[g(i\varepsilon,j\varepsilon,k\varepsilon)+3 g(i\varepsilon,j\varepsilon,k\varepsilon)^{2}+
      2 g(i\varepsilon,j\varepsilon,k\varepsilon)^{3}\right]\right\}  + \mathcal{O}(\varepsilon^{4})
\end{multline}
%% Wrong power of $\varepsilon$ below!!

Here and below we will use the notation:
\begin{equation}
  \label{eq:5}
  g(x,y,z)=\frac{1}{e^{x+y+z}-1}
\end{equation}

This expression looks like a combination of Riemann sums for some integrals. Note that these sums
are finite, but some of the corresponding integrals are divergent.

>>>>

\section{The computation of the free energy asymptotic expansion}
\label{sec:free-energy-scaling}
First we substitute MacMahon formula \eqref{eq:12} into the free energy definition \eqref{eq:17} and
obtain
\begin{equation}
  \label{eq:20}
    %-\frac{f}{T}
    f=-\frac{1}{V}\ln Z =- \sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \frac{1}{V}
  \ln\left(\frac{1-e^{-\varepsilon (i+j+k)} e^{\varepsilon}}{1-e^{-\varepsilon (i+j+k)} e^{2\varepsilon}}\right)
\end{equation}
Here and below we will use the notation:
\begin{equation}
  \label{eq:5}
  g(x,y,z)=\frac{1}{e^{x+y+z}-1}
\end{equation}

Omitting the arguments of $g(i\varepsilon,j\varepsilon,k\varepsilon)$ we thus can rewrite the free energy as
\begin{equation}
  \label{eq:2}
  f=-\frac{1}{V}\sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \left[\ln\left(1+g-g e^{\varepsilon}\right)-\ln\left(1+g-g e^{2\varepsilon}\right)\right]
\end{equation}
We have $V=2(MN+MK+NK)=2(ab+bc+ca)\varepsilon^{-2}$ and $g\sim \varepsilon^{-1}$ as
$\varepsilon\to 0$. Let us rewrite the logarithms as a series in $\varepsilon, g$:
\begin{equation}
  \label{eq:25}
  f=-\frac{1}{2(ab+bc+ca)}\sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \varepsilon^{2}
  \left[-\sum_{l=1}^{\infty}\frac{1}{l}\left(\sum_{m=1}^{\infty}\frac{g\varepsilon^{m}}{m!}\right)^{l}+
    \sum_{l=1}^{\infty}\frac{1}{l}\left(\sum_{m=1}^{\infty}\frac{2^{m} g\varepsilon^{m}}{m!}\right)^{l}\right]
\end{equation}
We are interested in the leading contributions in $\varepsilon$. Since $g\sim \varepsilon^{-1}$,
they come from the terms $g^{l}\varepsilon^{l}$ in the square brackets. The coefficients before
these terms are $\frac{2^{l}-1}{l}$:
\begin{equation}
  \label{eq:36}
  f\approx-\frac{1}{2(ab+bc+ca)}\sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \varepsilon^{2} \left[g+\sum_{l=2}^{\infty}\frac{2^{l}-1}{l}g^{l}\varepsilon^{l}\right]
\end{equation}
We separated the first term since it plays a major role in the following computations.

%% Since $g\sim \varepsilon^{-1}$, the first two terms in $\varepsilon$ are
%% \begin{multline}
%% \label{eq:35}
%%   %\frac{f}{T}
%%   f\approx-\frac{1}{2(ab+bc+ca)} \sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K}\left\{
%%     \varepsilon^{3} g(i\varepsilon,j\varepsilon,k\varepsilon)+
%%     \frac{3}{2}\varepsilon^{4}\left[g(i\varepsilon,j\varepsilon,k\varepsilon)+g(i\varepsilon,j\varepsilon,k\varepsilon)^{2}\right]\right\}\\
%%   +\mathcal{O}(\varepsilon^{2})
%% \end{multline}
%% Contribution of the order $\varepsilon^{2}$ come from the terms $g^{l}\varepsilon^{l}$ from the square brackets of the equation \eqref{eq:25}. The coefficients before these terms are $\frac{2^{l}-1}{l}$
%% 
%% 
%% \begin{multline}
%%   \label{eq:4}
%%   %\frac{f}{T}
%%   f=- \sum_{i=1}^{M} \sum_{j=1}^{N} \sum_{k=1}^{K} \frac{1}{V}\left\{
%%     \varepsilon g(i\varepsilon,j\varepsilon,k\varepsilon)+
%%     \frac{3}{2}\varepsilon^{2}\left[g(i\varepsilon,j\varepsilon,k\varepsilon)+g(i\varepsilon,j\varepsilon,k\varepsilon)^{2}\right]\right.\\
%%   \left.+\frac{7}{6}\varepsilon^{3}\left[g(i\varepsilon,j\varepsilon,k\varepsilon)+3 g(i\varepsilon,j\varepsilon,k\varepsilon)^{2}+
%%       2 g(i\varepsilon,j\varepsilon,k\varepsilon)^{3}\right]\right\}  + \mathcal{O}(\varepsilon^{4})
%% \end{multline}
%% %% Wrong power of $\varepsilon$ below!!
%% 
%% 
This expression looks like a combination of Riemann sums for some integrals. Note that these sums
are finite for each $l$, but some of the corresponding integrals are divergent.
